---
title: "course-notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How do you install a package from CRAN?
# How do you install a package from GitHub
# How does installing a package change your
# computer?

# What is a library?
# Where's your default library?
# Which libraries are searched for packages?
# How many packages do you have installed?

```{r}
.Library
.libPaths()
installed.packages()
```

When you install a package, a file will be downloaded and unzipped inside the library directory.

```{r}
lista <- installed.packages() %>%
  as_tibble() %>%
  select(1:4)

lista %>%
  count(Priority)

lista %>%
  filter(...)
```

# What packages do you currently have attached?
# What libraries did they come from?

```{r}
search()
searchpaths()
```


# How does library() affect these functions?

```{r}
library(tidyverse)
search()
searchpaths()
```

quando você escreve o nome de uma função, R vai procurar em toda a lista do `search()`, na ordem. Na Global Environment, depois nos outros pacotes. E `searchpaths()` mostra onde estão esses pacotes nos diretórios.

Base package is the lowest level package.

R --vanilla

no terminal

when you call library for a package, R will add this package to the searchpath.

## The Whole Game

A package is a set of
conventions that
(with the right tools)
makes your life easier 

```{r}
usethis::create_package("~/foofactors")
```

In a installed package directory, in the R folder, you won't find any R code. It's kinda compilled.

`usethis::use_r("fbind")` creates a function named "fbind", using the convention.

factors under the hood are simply integer.

devtools::load_all() ~~ install.packages() + library()

não usar packages de análise no Rprofile, pq prejudica reprodutibilidade.

document() ~~ control + shift + I

## tests

a bit like double-entry bookkeeping


