---
title: "Notes -- Working with Data in the Tidyverse"
output:
  html_document:
    df_print: paged
---

## Explore your data

### Import Data

*Rectangular Data:* columns hold variables.

Column values in the same row should correspond to the same observation.

Package `readr`.

### Reading CSV files

```{r}
library(readr)
?read_csv
```

When reading the manual, you'll notice that some arguments have a default value assigned.

> Default values are sensible. So it's good practice to start experimenting with arguments that don't have default values first.

```{r, eval=FALSE}
# Load readr
library(readr)

# Create bakeoff but skip first row
bakeoff <- read_csv("bakeoff.csv", skip = 1)

# Print bakeoff
bakeoff
```

Our original import had 3 problems: we had 1 extra observation, our first row stored the true variable names, and all column types were characters. Adding `skip = 1` worked because the default argument for `col_names` is `TRUE`.

### Assign Missing Values

The `read_csv()` function also has an `na` argument, which allows you to specify value(s) that represent missing values in your data. The default values for the `na` argument are `c("", "NA")`, so both are recoded as missing (`NA`) in R. When you read in data, you can add additional values like the string "`UNKNOWN`" to a vector of missing values using the `c()` function to combine multiple values into a single vector.

```{r, eval=FALSE}
# Load dplyr
library(dplyr)

# Filter rows where showstopper is UNKNOWN 
bakeoff %>% 
    filter(showstopper == "UNKNOWN")

# Edit to add list of missing values
bakeoff <- read_csv("bakeoff.csv", skip = 1,
                    na = c("", "NA", "UNKNOWN"))

# Filter rows where showstopper is NA 
bakeoff %>% filter(is.na(showstopper))
```

### Know your data

`glimpse`
`skim`, from `skimr`

```{r}
library(skimr)
glimpse(starwars)
skim(starwars)
```

You can combine skim() with other functions in a sequence using the pipe (%>%) operator. For example, you could use other dplyr functions like group_by first, then use skim() by adding a line after the final pipe.

This will produce summary statistics for each value of the grouping variable. Let's practice this!

First filter() to extract only those observations where us_season is not missing, then pipe that to skim(). Recall that ! means "not", so !is.na() creates the right logical condition.

```{r}
# Load skimr
library(skimr)

# Edit to filter, group by, and skim
bakeoff %>% 
  filter(!is.na(us_season)) %>% 
  group_by(us_season) %>% 
  skim()
```



